---
title: "cohortRand, a randomization package for cohort randomization"
author:
- name: Kai Xia
  affiliation: University of North Carolina at Chapel Hill
output:
  html_document:
    df_print: paged
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Introduction

For Randomized Controlled Trial (RCT) of intervention studies, the participants are usually randomized as group or cohort to a few number of intervention arms. In this case, instead of randomizing subject by subject, the RCT randomization is done by group or cohort. Minimization method tends to minimize the differences among groups while controlling for a number of prognostic factors. cohortRand is designed to provide user friendly R implementation to perform cohort randomization using existing miniRand R package. It also provides summary statistics to track how each cohort has been randomized.


## Key features
1. Randomize RCT subject by cohort using minization method
2. Provide summary statistics for randomized subjets

## Install GitHub Version
To install `cohortRand` directly from GitHub, run

```r
if(!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("argossy/cohortRand@main")
```

The package includes reference manual, sample data and a Vignette.

## Basic Usage
### Randomizing cohort 1
```{r echo=TRUE,results='asis',include=TRUE, cache=FALSE,message=FALSE, warning=FALSE, eval=TRUE}
## simulation code for random assignments

## add cohort by cohort in each chunk, then save the output by the date
## Cohort 1

library(devtools)
library(gtsummary)
library(knitr)
library(Minirand) # minization method


dir_pack = '..' # if on PC and linux if folder is correctly placed
load_all(sprintf('%s/cohortRand',dir_pack))



### run example



################################################################################
# randomizing cohort 1
################################################################################

## cohort number
co_id = 1
method='Range'

fileout = sprintf('random_sequence_cohort_%s.csv', co_id)
set.seed(2023 + co_id)
ntrt <- 3 # number of treatment types
#trtseq <- 1:ntrt # treatment assignment number
rand.mat <- NULL # randomization matrix


## simulating the information of this cohort
  nsample <- round(runif(1,18,22)) # sample size of this cohort
  ratio <- rep(1,ntrt) # assignment ratio
  
  # simulating two covariates
  c1 <- sample(c('M', 'F'), nsample, replace = TRUE, prob = c(0.4, 0.6))
  c2 <- sample(c('H','L'), nsample, replace = TRUE, prob = c(0.3, 0.7))
  covmat <- cbind(c1, c2) # generate the matrix of covariate factors for the subjects
  # label of the covariates. start with 'CV' so that it can be identified as covariate (strata)
  colnames(covmat) = c("CV.Gender", "CV.Risk")
  
  # add ID for simulated data
  id_start = 1
  id_end = id_start + nsample - 1
  ID = paste('ID_', id_start:id_end, sep='')
  covmat <- data.frame(ID, covmat)
  
  sum_cov(covmat )
  ## create input matrix with ID, covariates and empty randomization
  rand.mat <- create_rand_mat(rand.mat, covmat)

  ## start cohort 1 randomization using minization method
  rand.mat <- cohortRand(rand.mat, ntrt = ntrt, ratio = ratio,method=method)
  write.csv(rand.mat, file = fileout,row.names = FALSE,quote=FALSE)
  

### output the randomization assigment table
sum_random(rand.mat)

```

### Randomizing cohort 2
```{r echo=TRUE,results='asis',include=TRUE, cache=FALSE,message=FALSE, warning=FALSE, eval=TRUE}
## simulation code for random assignments

## add cohort by cohort in each chunk, then save the output by the date
## Cohort 2

## cohort ID
co_id = 2

fileout = sprintf('random_sequence_cohort_%s.csv', co_id)

### run example


set.seed(2023 + co_id)
#ntrt <- 2
#trtseq <- 1:ntrt
#rand.mat <- read.csv(filein,header = TRUE)

   nsample <- round(runif(1,18,22))
  
  ratio <- rep(1,ntrt)
  c1 <- sample(c('M', 'F'), nsample, replace = TRUE, prob = c(0.4, 0.6))
  c2 <- sample(c('H','L'), nsample, replace = TRUE, prob = c(0.3, 0.7))
  covmat <- cbind(c1, c2) # generate the matrix of covariate factors for the subjects
  # label of the covariates
  colnames(covmat) = c("CV.Gender", "CV.Risk")
  
  ## create input matrix with both covartes and randomization
  id_start = dim(rand.mat)[1] + 1
  id_end = id_start + nsample - 1
  ID = paste('ID_', id_start:id_end, sep='')
  covmat <- data.frame(ID, covmat)
  sum_cov(covmat )
  
  rand.mat <- create_rand_mat(rand.mat, covmat)
  
  ## start cohort 1 randomization using minization method
  rand.mat <- cohortRand(rand.mat, ntrt = ntrt, ratio = ratio,method=method)
  write.csv(rand.mat, file = fileout,row.names = FALSE,quote=FALSE)
  
  ### output summary of randomization table
  sum_random(rand.mat)


```



```{r echo=FALSE,results='asis',include=TRUE, cache=FALSE,message=FALSE, warning=FALSE, eval=TRUE}
## simulation code for random assignments

## add cohort by cohort in each chunk, then save the output by the date
## Cohort 3

library(devtools)
library(gtsummary)
library(knitr)
library(Minirand)

## cohort ID
co_id = 3

# dir_pack = '..' # if on PC and linux if folder is correctly placed
# load_all(sprintf('%s/cohortRand',dir_pack))

filein =  sprintf('random_sequence_cohort_%s.csv', co_id-1)
fileout = sprintf('random_sequence_cohort_%s.csv', co_id)

### run example


set.seed(2023 + co_id)
# ntrt <- 2
# trtseq <- 1:ntrt
#rand.mat <- read.csv(filein,header = TRUE)

  nsample <- round(runif(1,18,22))
  
  ratio <- rep(1,ntrt)
  c1 <- sample(c('M', 'F'), nsample, replace = TRUE, prob = c(0.4, 0.6))
  c2 <- sample(c('H','L'), nsample, replace = TRUE, prob = c(0.3, 0.7))
  covmat <- cbind(c1, c2) # generate the matrix of covariate factors for the subjects
  # label of the covariates
  colnames(covmat) = c("CV.Gender", "CV.Risk")
  
  ## create input matrix with both covartes and randomization
  id_start = dim(rand.mat)[1] + 1
  id_end = id_start + nsample - 1
  ID = paste('ID_', id_start:id_end, sep='')
  covmat <- data.frame(ID, covmat)
  
  sum_cov(covmat )
  
  rand.mat <- create_rand_mat(rand.mat, covmat)
  
  ## start cohort 1 randomization using minization method
  rand.mat <- cohortRand(rand.mat, ntrt = ntrt,ratio=ratio,method=method)
  write.csv(rand.mat, file = fileout,row.names = FALSE,quote=FALSE)

  ### output summary of randomization table
  sum_random(rand.mat)



```






